---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: html_notebook
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Multi-category dot-density maps often work well when the categories cluster geographically. Recent immigrants by
country of origin work well for this.
## Data
First we grab the immigrant data via [cancensus](https://github.com/mountainMath/cancensus), making use of the [CensusMapper API tool](https://censusmapper.ca/api/CA11) to select the regions and variables we need.
```{r, message=FALSE, warning=FALSE}
#devtools::install_github("mountainmath/cancensus")
library(cancensus)
library(dotdensity)
# options(cancensus.api_key='your_api_key')
regions=list(CMA="59933")
vectors=c("v_CA11N_265","v_CA11N_268","v_CA11N_304","v_CA11N_334","v_CA11N_373","v_CA11N_376","v_CA11N_379","v_CA11N_382")
```

We choose the categories and colours we want to map and define a convenience function to rename the variables and compute the qantities for the other asian countries that we don't break out.
```{r}
categories=c("Americas","Europe","Africa","Philippines","China","India","Other Asian Countries")
colors=c("#7a0177", "#3333cc", "#ff00ff", "#00ffff", "#ff1a1c", "#4dff4a", "#ffff33")

prep_data <- function(geo){
  data <- geo@data %>% replace(is.na(.), 0)
  data <- rename(data,
    total=v_CA11N_265,
    Americas=v_CA11N_268,                                  
    Europe=v_CA11N_304,
    Africa=v_CA11N_334,
    Philippines=v_CA11N_376,
    China=v_CA11N_379,
    India=v_CA11N_382)
  data <- mutate(data,`Other Asian Countries` = v_CA11N_373-Philippines-China-India)
  geo@data <- data
  return(geo)
}
```


Next we grab the data via `cancensus`,
```{r, echo=TRUE, message=FALSE, warning=FALSE}
data_csd=get_census(dataset = 'CA11', regions=regions,vectors=vectors,geo_format='sp',labels='short',level='CSD') %>% prep_data
data_ct=get_census(dataset = 'CA11', regions=regions,vectors=vectors,geo_format='sp',labels='short',level='CT') %>% prep_data
data_da=get_census(dataset = 'CA11', regions=regions,vectors=vectors,geo_format='sp',labels='short',level='DA') %>% prep_data
data_db=get_census(dataset = 'CA11', regions=regions,geo_format='sp',labels='short',level='DB')
```

which we then re-aggregate to make sure we don't miss overall counts due to privacy cutoffs distribute them
proportionally among the population.
```{r}
data_ct@data <- dot_density.proportional_re_aggregate(data=data_ct@data,parent_data=data_csd@data,geo_match=setNames("GeoUID","CSD_UID"),categories=categories,base="Population")
data_da@data <- dot_density.proportional_re_aggregate(data=data_da@data,parent_data=data_ct@data,geo_match=setNames("GeoUID","CT_UID"),categories=categories,base="Population")
data_db@data <- dot_density.proportional_re_aggregate(data=data_db@data,parent_data=data_da@data,geo_match=setNames("GeoUID","DA_UID"),categories=categories,base="Population")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
bg_color="#111111"
base_color="#333333"
text_color="#eeeeee"
theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       panel.grid.major = element_blank(),
                       panel.background = element_rect(fill = bg_color, colour = NA),
                       plot.background = element_rect(fill=bg_color, size=1,linetype="solid",color=text_color),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       plot.title = element_text(size=80,hjust = 0.5, color=text_color),
                       plot.subtitle = element_text(size=60,hjust = 0.5, color=text_color),
                       plot.caption = element_text(size=25, color=text_color),
                       legend.title = element_text(size=40, color=text_color),
                       legend.text = element_text(size=40, color=text_color),
                       legend.background = element_rect(fill=bg_color, size=1,linetype="solid",color=bg_color),
                       legend.key = element_rect(fill = bg_color,color = bg_color),
                       legend.key.width = unit(3, 'lines'),
                       legend.position = "bottom"))

basemap <-   ggplot(data_csd) +
    geom_polygon(aes(long, lat, group = group), fill = base_color, size=0.1, color = 'grey') +
    guides(colour = guide_legend(nrow=1,override.aes = list(size=15))) +
    coord_map(projection="lambert", lat0=49, lat=49.4) +
    theme_opts

```


##Map
All that's left to do is to covert our re-aggregated block-level data to dots, using the `dot_density.compute_dots`
function from the [`dotdensity` package]() and feed it into the `dot_density.dots_map` function to add them to
our basemap.
```{r, fig.height=10, fig.width=13, message=TRUE, warning=TRUE}
# 1 dot = 5 immigrants
scale=5


dots.db <- dot_density.compute_dots(geo_data = data_db, categories = categories, scale=scale)
basemap +
  # zoom in a bit
  coord_fixed(xlim=c(-123.29,-122.6), ylim=c(49.02,49.35), ratio = 1/cos(49.2/180*pi)) +
# shade unpopulated blocks
#  geom_polygon(data=data_db[data_db$Population<=5,], 
#                        aes(long, lat, group = group), 
#                        fill = "#222222", size=0.1, 
#                        color = "#222222") +
  scale_colour_manual(values = colors) +
  labs(color = "",
                title="Immigrants 2006 - 2011",
                caption="Source: StatCan Census 2016 via cancensus & CensusMapper.ca",
                subtitle = paste0("1 dot = ",scale," people")) + 
  dot_density.dots_map(dots=dots.db,alpha=0.75,size=0.25)

# save image for later
ggsave('../images/recent_immigrants.png',width=26,height=26)
```

